---
title: "p8105_hw1_hp2661"
author: "Huizhong Peng"
date: "2024-09-17"
output: github_document
---

```{r include=FALSE}
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
```

# Problem 1

Download the package in the console and load the dataset.

```{r results='hide'}
data("penguins", package = "palmerpenguins")
head(penguins)
summary(penguins)

nrow(penguins)
ncol(penguins)
mean(penguins$flipper_length_mm, na.rm = T)
```

The penguins dataset contains 344 rows and 8 columns, including variables 
"species", "island", "bill_length_mm", "bill_depth_mm", "flipper_length_mm", 
"body_mass_g", "sex" and "year". 

The species includes "Adelie", "Chinstrap" and "Gentoo". 

The mean flipper length is 200.9 mm.

Then create a scatter plot.

```{r warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) + 
  geom_point() +
  labs(title = "Scatterplot of Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)")
ggsave("hw1_scatter_plot.pdf", height = 4, width = 6)
```


# Problem 2

Create a dataframe.

```{r}
set.seed(1234)

dat = tibble(
  vec_numeric   = rnorm(10, mean = 0, sd = 1), 
  vec_logical   = vec_numeric >0,
  vec_character = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor    = factor(c("low", "moderate", "high", "low", "moderate", "high", 
                          "low", "moderate", "high", "high"))
)
print(dat)
```


```{r, results='hide', warning=FALSE}
dat |> pull(vec_numeric) |> mean()
dat |> pull(vec_logical) |> mean()
dat |> pull(vec_character) |> mean()
dat |> pull(vec_factor) |> mean()
```

If taking the mean, the numeric vector and logical vector work. 
The character vector and factor vector don't work, returning NA.


Convert the logical, character, and factor variables to numeric variables.

```{r, x-type, eval=FALSE}
num_log = 
  dat |> pull(vec_logical) |> as.numeric()
num_cha = 
  dat |> pull(vec_character) |> as.numeric()
num_fac = 
  dat |> pull(vec_factor) |> as.numeric()

num_log
num_cha
num_fac
```

The logical variables are successfully converted to numeric variables. "TRUE" 
becomes 1 and "FALSE" becomes 0. This explains why taking the mean of a logical 
vector works, and the mean reflects the proportion of "TRUE" in this vector.

The conversion of character variables fail and it returns NA. This is because
characters can't be interpreted as numbers, and it explains why taking the mean 
of a character vector fails.

The factor variables are successfully converted to numeric variables because 
each level corresponds to a number. "low" becomes 2, "moderate" becomes 3 and 
"high" becomes 1. But it has no meaning in practice so mean() can't take the 
mean of a factor vector.











